---
title: "DSerna_LiveSession10Assignment"
author: "Daniel Serna"
date: "July 22, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
```{r loadDataFrame}
load("N-MHSS-2015-DS0001-data-r.rda")
if(!require(sqldf)) install.packages("sqldf")
```

### Question 1b
```{r stateAbbreviations}
#create a vector of state abbreviations with whitespace trimmed..
stateAbbreviations <-trimws(as.vector(unique(mh2015_puf$LST)), which = c("both"))

#sort vector alphabetically.
stateAbbreviations <- sort(stateAbbreviations)

#display state abbreviations.
stateAbbreviations
```

### Question 1c
```{r hospitalCount}
#create a vector for states to exclude
statesToExclude <- c("AK", "AS", "GU", "HI", "PR", "VI")

#exclude everything but mainland states
statesFiltered <- setdiff(stateAbbreviations, statesToExclude)

#filter dataframe for only mainlaind states
mhdfFiltered <- mh2015_puf[trimws(mh2015_puf$LST, which = c("both")) %in% statesFiltered & mh2015_puf$FACILITYTYPE =="Veterans Administration medical center (VAMC) or other VA health care facility",]

#load sqldf library to handle data queries
library(sqldf)

#create a dataframe that lists the VA hospital count by state
vaCount<- sqldf("select LST as State, Count(*) as VACount from mhdfFiltered group by LST")
vaCount
```
### Question 1d
```{r plotHospitalCount}
#load ggplot2 library
library(ggplot2)

#plot count by state, remove legend, and flip orientation of x-axis.
vaCountPlot <- ggplot(data = vaCount, aes(x = vaCount$State, y = vaCount$VACount, fill = vaCount$State)) + geom_bar(stat="identity") + ggtitle("VA Hospital count by State") +
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle=90)) + 
  ylab("Hospital Count") + 
  xlab("State Abbreviation") + labs(fill = "State") + guides(fill=FALSE)
vaCountPlot
```

## Question 2

### Question 2a
```{r importStateSize}
stateSize <- read.csv("statesize.csv")
paste(vaCount$State)
#The issue is there is whitespace in the state abbreviation field.
```

### Question 2b
```{r rectifyWhiteSpaceAndMerge}
#trim off the white space
vaCount$State <- trimws(vaCount$State, which = c("both"))

#display rectified state abbreviation.
vaCount$State

#rename column in preparation for merge
colnames(vaCount) <- c("Abbrev", "VACount")

#merge data frames
mergedDF <- merge(vaCount, stateSize, by="Abbrev")
#display head of merged data frame
head(mergedDF)
```

### Question 2c
```{r countHospitalsPer1000SQMiles}
#calculate the count of hospitals per 1k square miles
mergedDF$HospitalsPer1kSQMiles <- (mergedDF$VACount / (mergedDF$SqMiles / 1000))

#display head of data frame with new column
head(mergedDF)
```

### Question 2d
```{r plotHospitalsPer1kSQMiles}
#plot count by 1k square miles, order bars in decreasing order, change default fill color, and flip orientation of x-axis.
va1kCountPlot <- ggplot(data = mergedDF, aes(x = reorder(mergedDF$Abbrev, -mergedDF$HospitalsPer1kSQMiles), y = mergedDF$HospitalsPer1kSQMiles, fill = mergedDF$Region)) + geom_bar(stat="identity") + scale_fill_manual("legend", values = c("Midwest" = "purple", "Northeast" = "orange", "South" = "red", "West" = "black")) + ggtitle("VA Hospital count per 1k square miles by state") +
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle=90)) + 
  ylab("VA hospital count per 1k square miles") + 
  xlab("State Abbreviation") + labs(fill = "Region")
va1kCountPlot
```

### Question 2e
It appears the Northeast has a disproportionately higher amount of VA hospitals per 1000 square miles when compared to the other regions. The West appears to have a disproportionately lower amount of VA hospitals per 1000 square miles when compared to the other regions. I would recommend we try to rectify the low number of VA hospitals in the West region. Perhaps we could funnel some funding from the Northeast region into the West to rectify the situation. I would also recommend we ensure we have accurate data before making budget priority changes.

## Github repo location
https://github.com/djserna/msds-6306-homework/tree/master/Unit10
