---
title: "DSerna_LiveSession11Assignment"
author: "Daniel Serna"
date: "July 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

### Question 1a

```{r euStockMarkets}
#display help file for EuStockMarkets data set.
#help("EuStockMarkets") commenting this out as we don't want to run this when knitting.

#pull the DAX index into a variable.
dax <- EuStockMarkets[,"DAX"]
```

### Question 1b
```{r plotDAX}
#create a plot with blue line and red indicator divding pre and post 1997 data.
plot(dax, xlab="Year", ylab="DAX Index", main="DAX index over time from 1991 to 1999", col="blue")
abline(v=1997, col="red")
```

### Question 1c
```{r decomposeTimeSeries}
#get a decompose variable for dax index.
decompose_dax <- decompose(dax, "multiplicative")

#plot the seasonal component.
plot(decompose_dax$seasonal, xlab="Year", ylab="DAX index seasonal component", main="DAX index seasonal component from 1991 to 1999", col="blue")
abline(v=1997, col="red")

#plot the trend componenet
plot(decompose_dax$trend, xlab="Year", ylab="DAX index trend component", main="DAX index trend component from 1991 to 1999", col="blue")
abline(v=1997, col="red")

#plot the random componenet
plot(decompose_dax$random, xlab="Year", ylab="DAX index random component", main="DAX index random component from 1991 to 1999", col="blue")
abline(v=1997, col="red")
```

## Question 2

### Question 2a
```{r maxTemp}
if (!require(fpp2)) install.packages("fpp2")
library(fpp2)
#help(maxtemp) commenting this out as we don't want to run this when knitting.
autoplot(maxtemp)
```

### Question 2b
```{r subsetMaxtemp}
#create variable to hold data from 1990 onward.
maxtempSubset <- window(maxtemp, start=1990)
```

### Question 2c
```{r sesPredict}
fitSES1 <- ses(maxtempSubset, h=5)
plot(maxtempSubset, xlab="Year", ylab="Max temperature", main="Max temperatures from 1990-2021", xlim=c(1990,2021))
lines(fitted(fitSES1), col="blue")
lines(fitSES1$mean, col="blue", type="o")
fitSES1$model$aicc
```
AICc of the SES fitted model is 141.5302

### Question 2d
```{r holtPredit}
fitHolt1 <- holt(maxtempSubset, initial="optimal", damped = TRUE, h=5)
plot(maxtempSubset, xlab="Year", ylab="Max temperature", main="Max temperatures from 1990-2021", xlim=c(1990,2021))
lines(fitted(fitHolt1), col="blue")
lines(fitHolt1$mean, col="blue", type="o")
fitHolt2 <- holt(maxtempSubset)
fitHolt1$model$aicc
```
AICc of the Holt fitted model is 142.2355

### Question 2e
The AICc for the SES fitted model is 141.5302 which is less than the AICc of the damped Holt fitted model of 142.2355. Lower AICc is better so the SES fitted model is better in this example.

## Question 3

### Question 3a
```{r readDataFrames}
if (!require(dygraphs)) install.packages("dygraphs")
library(dygraphs)
olli <- read.csv("Unit11TimeSeries_Ollivander.csv", header = FALSE)
colnames(olli) <- c("Date", "OlliWandsSold")
greg <- read.csv("Unit11TimeSeries_Gregorovitch.csv", header = FALSE)
colnames(greg) <- c("Date", "GregWandsSold")
```
### Question 3b
```{r convertDate}
#convert date columns into date class
olli$Date <- as.Date(olli$Date, format = "%m/%d/%Y")
greg$Date <- as.Date(greg$Date, format = "%m/%d/%Y")
```

### Question 3c
```{r convertXTS}
#load xts library
library(xts)
#convert olli data frame to XTS object
olliXTS <- xts(olli, order.by = olli$Date)

#convert greg data frame to XTS object
gregXTS <- xts(greg, order.by = greg$Date)
```

### Question 3d
```{r dygraph}
#merge the two xts objects
merged <- merge(olliXTS, gregXTS)

#create a dygraph with shading, range selector, and highlighting.
dygraph(merged, xlab = "Date", ylab = "Wands Sold", main = "Number of wands sold by year") %>% dySeries("OlliWandsSold", color = "red") %>% dySeries("GregWandsSold", color = "green") %>% dyShading(from = "1995-1-1", to = "1999-1-1") %>% dyHighlight(highlightSeriesOpts = list(strokeWidth = 3)) %>% 
  dyRangeSelector(height = 100, dateWindow = c("1970-1-1","2017-1-1")) 
```

## GitHub repo
https://github.com/djserna/msds-6306-homework/tree/master/Unit11